---
title: "Lab, Part 2: `Oscars` Data Analysis"
# author: Isabella Benabaye
# date: 4/12/2020
output: github_document
---

## Import libraries and data

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)

oscars <- read_csv('https://raw.githubusercontent.com/Cal-Poly-Advanced-R/Lab_1/master/Oscars-demographics-DFE.csv')
```

## Tidy data
Upon checking the data, we can see there aresome inconsistencies.  

For `date_of_birth`:   

* One person has ` [1]` at the end of their birthday   
* Some of the dates have years that are 2 digits, some 4. If we simply use `dmy()`, the years before 1969 will be put in the 21st century instead of the 20th.   
* There is one person who only has a birth year listed.    

```{r age function, message=FALSE, warning=FALSE, include=FALSE}
age_years <- function(earlier, later)
{
    lt <- data.frame(earlier, later)
    age <- as.numeric(format(lt[,2],format="%Y")) - as.numeric(format(lt[,1],format="%Y"))
    
    dayOnLaterYear <- ifelse(format(lt[,1],format="%m-%d")!="02-29",
                             as.Date(paste(format(lt[,2],format="%Y"),"-",format(lt[,1],format="%m-%d"),sep="")),
                             ifelse(as.numeric(format(later,format="%Y")) %% 400 == 0 | as.numeric(format(later,format="%Y")) %% 100 != 0 & as.numeric(format(later,format="%Y")) %% 4 == 0,
                                    as.Date(paste(format(lt[,2],format="%Y"),"-",format(lt[,1],format="%m-%d"),sep="")),
                                    as.Date(paste(format(lt[,2],format="%Y"),"-","02-28",sep=""))))
    
    age[which(dayOnLaterYear > lt$later)] <- age[which(dayOnLaterYear > lt$later)] - 1
    
    age
}
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
oscars$date_of_birth <- str_replace_all(oscars$date_of_birth," [1]","")
oscars <- oscars %>% 
  # fix the date of birth
  mutate(date_of_birth = if_else(str_length(date_of_birth) > 11, str_sub(date_of_birth,1,11), date_of_birth)) %>%  ## remove "[1]" from one of the obs
  separate(date_of_birth, c("day","month","birthyear"), sep = "-", remove = FALSE) %>%   ## separate birthday components
  mutate(birthyear = case_when(str_length(birthyear) == 2 ~ paste0("19",birthyear),  ## add the century to the birth year
                               str_length(day) == 4 ~ day,  ## address the obs with only the year
                               TRUE ~ birthyear),
         dob = ymd(paste(birthyear,month,day,sep = " ")),  ## get the formatted birthday
         award_dt = ymd(paste(year_of_award,"Feb","1", sep = " ")),
         ## calculate the person's age using the function age_years() from https://raw.githubusercontent.com/nzcoops/r-code/master/age_function.R
         age = age_years(as.Date(dob), as.Date(award_dt)),
         gender = case_when(award %in% c("Best Actor", "Best Supporting Actor") ~ "M",
                            award %in% c("Best Actress", "Best Supporting Actress") ~ "F",
                            TRUE ~ NA_character_)) 

```

## Warm-ups

1. Which movie(s) won the most unique "Big 5" awards?

```{r paged.print=FALSE}
most_awards <- oscars %>% 
  select(movie,award) %>% 
  distinct() %>%   ## make sure there are no duplicates
  group_by(movie) %>% 
  count(name = "awards") %>%  ## count the number of awards per movie
  ungroup() %>% 
  filter(awards == max(awards))  ## only get the top movies with the most unique awards
```
```{r echo=FALSE}
knitr::kable(most_awards)
```

2. Of all actresses who have won the Best Actress award, what are is the most common first name?

```{r paged.print=TRUE}
common_name <- oscars %>% 
  filter(award == "Best Actress") %>% 
  distinct(person)%>%  ## make sure each actress is only counted once
  mutate(first_name = word(person,1)) %>%  ## get their first names
  group_by(first_name) %>% 
  count() %>%  ## count the number of women with the same first name
  ungroup() %>% 
  filter(n == max(n)) ## only output the most common
```
```{r echo=FALSE}
knitr::kable(common_name)
```



3. What US State, or non-US country, has produced the most Oscar winners (for the awards in this dataset)?

```{r}
common_birthplace <- oscars %>% 
  select(person, birthplace) %>% 
  distinct() %>% 
  group_by(birthplace) %>%
  count() %>% 
  ungroup() %>% 
  filter(n == max(n))
```
```{r echo=FALSE}
knitr::kable(common_birthplace)
```

The most common birthplace is New York, since New York City has produced the most Oscar winners.

## Age and Gender

